Constraining the Interiors of Asteroids Through Close Encounters

### 2022-04-14 paper/appendices.tex
Ready to stage new sigma scan

\documentclass[fleqn,usenatbib]{mnras}

\usepackage[T1]{fontenc}

\DeclareRobustCommand{\VAN}[3]{#2}
\let\VANthebibliography\thebibliography
\def\thebibliography{\DeclareRobustCommand{\VAN}[3]{##3}\VANthebibliography}

%%%%% AUTHORS - PLACE YOUR OWN PACKAGES HERE %%%%%

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage{xcolor}
\usepackage{tikz}
\usepackage{colortbl}
\usepackage{dsfont}
\usepackage{bm}

% For the sideways figure
\usepackage{pdflscape}
\usepackage{afterpage}

\usetikzlibrary{arrows.meta}

@@

%%%%% AUTHORS - PLACE YOUR OWN COMMANDS HERE %%%%%
\newcommand{\jtd}[1]{ {\bf{\color{red} JTD: #1}} }
\newcommand{\unit}[1]{\bm{\hat{#1}}}

\newcommand{\parens}[1]{\left( #1 \right)}
\newcommand{\brackets}[1]{\left[ #1 \right]}
\newcommand{\quat}[1]{\widetilde{\bm{#1}}}

% Please keep new commands to a minimum, and use \newcommand not \def to avoid
% overwriting existing commands. Example:
%\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared

%%%%%%%%%%%%%%%%%%% TITLE PAGE %%%%%%%%%%%%%%%%%%%

\title[Flyby Constraints on Asteroids Interiors]{Constraining the Interiors of Asteroids Through Close Encounters}

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[Jack T. Dinsmore, Julien de Wit]{
Jack T. Dinsmore,$^{1}$\thanks{E-mail: jtdinsmo@mit.edu}
Julien de Wit$^{2}$
\\
% List of institutions
$^{1}$Department of Physics, Massachusetts Institute of Technology\\
$^{2}$Department of Earth, Atmospheric, and Planetary Science, Massachusetts Institute of Technology
}

% These dates will be filled out by the publisher
\date{Accepted XXX. Received YYY; in original form ZZZ}

% Enter the current year, for the copyright statements etc.
\pubyear{2022}

% Don't change these lines
\begin{document}
\label{firstpage}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\maketitle

% Abstract of the paper
\begin{abstract}
  We investigate the degree to which asteroid interior density distributions can be extracted from rotational velocity data gathered during a close encounter. We derive the equations of motion for an rigid asteroid's orientation and angular velocity to arbitrary order and use them to generate synthetic rotational velocity data for a representative asteroid on a close Earth encounter. Using Markov Chain Monte Carlo fits, we re-extract the density moments of the asteroid in a wide range of scenarios to measure the degree to which best-fitting parameter precision is affected. Specifically, we use many injection-retrieval tests to study fit precision's dependence on the asteroid's moment of inertia, observational precision and cadence, orbital parameters, and initial spin pole direction as well as the quantity of near-pericentre data and the central body oblateness. Finally, we discuss the degeneracy between the density moments and the actual density distribution and propose three models to construct a representative density distribution from fit results.
\end{abstract}

% Select between one and six entries from the list of approved keywords.
% Don't make up new ones.
\begin{keywords}
  minor planets, asteroids: general -- methods: data analysis
\end{keywords}

%%%%%%%%%%%%%%%%% BODY OF PAPER %%%%%%%%%%%%%%%%%%

\section{Introduction}

Over the past twenty years, the increase in quantity and quality of high-sensitivity, all-sky surveys has prompted the discovery of numerous asteroids. Such advances have been made via ground-based surveys such as the Catalina Sky Survey \cite{larson1998catalina}, Pan-STARRS \cite{kaiser2002pan}, and the Lincoln Near-Earth Asteroid Research project (LINEAR) \cite{stokes2000lincoln}, as well as space-based instruments such as the Wide-field Infrared Survey Explorer (WISE) mission \cite{wright2010wide}. Many of these asteroids are relatively small, but some are kilometre-sized and a few are predicted to encounter Earth or other planets closely in the near future. Light curve analysis has been performed on many of these new asteroids to study their surface properties, but analyses that are sensitive to the interior structure of the asteroid are rarer.

Theoretical work has demonstrated that tidal torque alters the angular velocity of an asteroid during a gravitational encounter with another body, and that this tidal torque is sensitive to the interior density distribution of the asteroid \cite{paul88, SCHEERES2000106, ashenberg07, BOUE2009750, HouMar2017}. This two-body generalized system has been studied to different orders and with several different methods. The angular velocity changes induced by tidal torque occur both in asteroid binaries \cite{Naidu_2015, Makarov2022ChaosOO}, and during planetary encounters \cite{RICHARDSON199847, scheeres2004evolution}. Such asteroids need not be tidally disrupted for their rotation state to be perturbed by tidal torque. Some examples of rotational velocity altered during an encounter have recently been observed, including for the 2013 encounter of (367943) Duende with Earth \cite{MOSKOVITZ2020113519, benson2020spin}, and have been predicted for the 2029 encounter of (99942) Apophis with Earth \cite{SCHEERES2005281, DEMARTINI201993}. The upcoming encounter of Apophis is of particular interest and much attention has been devoted to accurate estimates of its orbital and physical properties \cite{yu2014numerical, hirabayashi2021finite,valvano2022apophis, Lee2022Apophis}, and in particular how its angular velocity will be affected by tidal torque \cite{souchay2014rotational, souchay2018changes}. Altered rotational states altered by Solar tidal torque have been excluded for 1I/2017 U1 `Oumuamua \cite{KWIECINSKI2018170}.

Such asteroid encounters have further been shown to reveal the physical properties of the encountering asteroid. Refs.~\cite{MOSKOVITZ2020113519, benson2020spin} extract best-fitting dimensions for a triaxial ellipsoid model of Duende from light-curve data, and Refs.~\cite{DESCAMPS2020113726} and \cite{BERTHIER2020113990} extract bulk densities and rotational periods for asteroid binaries (3905) Doppler and (617) Patroclus respectively.

More encounter candidates are likely to be discovered in the future --- for Earth or other planets --- potentially by the James Webb Space Telescope (JWST) or Large-aperture Synoptic Survey Telescope (LSST), and general methods for assessing whether an asteroid's encounter properties are conducive to the extraction of physical asteroid characteristics from light-curve data is necessary. However, such general encounters are not very well studied, since much of previous research has focused on specific encounters.

Furthermore, asteroid properties which appear in the beyond-leading-order terms of tidal torque have not yet been extracted, and more research is needed to study in what cases these effects are observable.

In this paper, we address these two topics by deriving a new equation of motion for an asteroid encounter system in section \ref{sec:methods}. This equation is simple and exact to arbitrary order, given assumptions reasonable for a planetary encounter (and many binary asteroid systems). It is also fast in that it allows volume integrals over the asteroid to be pre-computed. We use this equation of motion to simulate a generic asteroid encounter of Earth, and extract physical characteristics of the asteroid (specifically, density moments) from synthetic data via a Markov Chain Monte Carlo (MCMC) fit, described in section \ref{sec:fit}. We measure the uncertainties in the posterior distribution of density moments as a function of various encounter parameters such as the observational uncertainty, cadence of observation, encounter orbital parameters, and more. We express which parameters most affect the precision of fit results in section \ref{sec:results}. Finally, we propose methods to extract full density distributions with uncertainties from these density moments in section \ref{sec:distros}. We test the behaviour of these methods, highlight their strengths and weaknesses, and suggest ways in which they may be altered to suit the needs of the user.

\section{Methods}
\label{sec:methods}

In this section, we describe our coordinates (section \ref{sec:coordinates}) for an encountering asteroid's position and orientation, and we parametrize its density distribution via its ``density moments'' (section \ref{sec:moments}). Then we derive an arbitrary-order equation for tidal torque (section \ref{sec:tidal-torque}), write the equations of motion for the system (section \ref{sec:eom}), and describe a simulation used to integrate the equation of motion and produce synthetic data of angular velocity over time (section \ref{sec:sim}).

@@

We define the density moments of an asteroid as
\begin{equation}
  K_{\ell m} = \frac{1}{\mu_\mathcal{A} a_\mathcal{A}^\ell} \int_\mathcal{A} d^3 r \rho_\mathcal{A}(\bm r) R_{\ell m}(\bm r).
  \label{eqn:klm}
\end{equation}
Note that these are complex. Here, $\mathcal{A}$ indicates the volume of the asteroid, $\mu_\mathcal{A}$ is the mass of the asteroid, $\rho_\mathcal{A}(\bm r)$ is the density distribution, and $a_\mathcal{A}$ is a real length scale defined (for reasons covered later) as 
\begin{equation}
  a_\mathcal{A}^2 = \frac{1}{\mu_\mathcal{A}} \int_\mathcal{A} d^3 r \rho_\mathcal{A}(\bm r) r^2.
  \label{eqn:am}
\end{equation}
This length scale can be thought of as akin to the radius of the asteroid (although a spherical, uniform-density asteroid has radius $a_\mathcal{A}\sqrt{5/3} $). Both equations \ref{eqn:klm} and \ref{eqn:am} should be computed in the body-fixed axes.

@@

Throughout this paper, we assume that the asteroid remains rigid throughout the encounter. We also assume no third-body perturbations from other Solar System objects. (Actually, third-body perturbing objects are allowed if they are closer to the central body's centre of mass than the asteroid perigee distance. Then, their density moments can be included in the density moments of the central body and this derivation can still be used.) For the sake of simplicity, we also assume that the density moments of the central body are known and do not evolve with time (i.e., the central body does not rotate).

@@

The tidal torque experienced by the asteroid (in the body-fixed frame) is given by
\begin{equation}
  \bm{\tau}(\bm u) = \int_\mathcal{A} d^3 u \rho_\mathcal{A}(\bm u) (\bm u \times (-\nabla_{\bm u} V(\bm u)))
\end{equation}
where $\rho_\mathcal{A}$ is the density distribution of the asteroid and $\mathcal{A}$ indicates the volume of the asteroid. Making use of one more identity concerning the derivatives of spherical harmonics:
\begin{equation}
  \begin{split}
  \bm u \times \nabla R_{\ell m}(\bm u)=&\frac{1}{2}\Big[(i\unit x - \unit y)(\ell-m+1)R_{\ell,m-1}(\bm u)\\
  &+(i\unit x+\unit y)(\ell+m+1)R_{\ell,m+1}(\bm u)\\
  & +2im\unit z R_{\ell m}(\bm u)\Big],
  \end{split}
\end{equation}
tidal torque can now be expressed as a function only of $J_{\ell m}$, $K_{\ell m}$, $a_\mathcal{A}$, $a_\mathcal{B}$, and the asteroid orientation and position. Explicitly,
\begin{equation}
  \begin{split}
  \bm \tau = & G\frac{\mu_\mathcal{A}\mu_\mathcal{B}}{2}\left[\sum_{\ell, m} a_\mathcal{B}^\ell J_{\ell m} \sum_{\ell', m'}a_\mathcal{A}^{\ell'}S^*_{\ell+\ell', m + m'} (\bm D) (-1)^{\ell'}\right.\\
  \times & \left.\sum_{m''=-\ell'}^{\ell'} \sqrt{\frac{(\ell'-m'')!(\ell'+m'')!}{(\ell'-m')!(\ell'+m')!}}  \mathcal{D}^{\ell'}_{m'm''}(\alpha, \beta, \gamma)^* \right. \\
  \times & \Big((i\unit x - \unit y)(\ell'-m''+1)K_{\ell',m''-1} \\
  & +(i\unit x+\unit y)(\ell'+m''+1)K_{\ell',m''+1}+2im''\unit z K_{\ell'm''}\Big) \Bigg].
  \end{split}
  \label{eqn:tidal-torque}
\end{equation}
Some $K_{\ell m}$ terms are included in this equation with $|m|>\ell$; these should all be taken to be zero.

@@

\subsection{Simulation design}
\label{sec:sim}

We built a custom simulation in C++ to produce angular velocity data (in the inertial frame, where it is observed from Earth) as a function of time. This simulation requires as initial data (1) the orbital parameters of the asteroid $r_p$ (perigee distance) and $v_\infty$ (hyperbolic excess velocity); (2) the cadence of angular velocity observation $\Delta t$; (3) the central body moments $J_{\ell m}$, mass $\mu_\mathcal{B}$, and radius $a_\mathcal{B}$; (4) the initial asteroid angular velocity in the inertial frame $\bm \Omega_0$; (5) the asteroid ``radius'' $a_\mathcal{A}$; and (6) the asteroid's density moments $K_{\ell m}$ and initial Euler angle $\gamma_0$. All parameters except (6) are assumed to be known to high accuracy. One can imagine that $a_\mathcal{A}$ is approximated by light-curve analysis, but if not, it is still necessary to fix $a_\mathcal{A}$ or else the values of $K_{\ell m}$ are degenerate with $a_\mathcal{A}$.

We further assume that the asteroid is initially not tumbling. Thus, the rotational velocity is aligned with a principal axis (assumed to be $\unit z$, which maximizes moment of inertia). This sets $\beta = 0$ and we can further choose $\alpha = 0$. Thus, only the Euler angle $\gamma_0$ is necessary to provide initial data for the simulation.

We begin our simulation at $D = 10 r_p$, with velocity predicted by Kepler's laws given by the orbital parameters. Since the leading order of the equations of motion is $\ell' = 2, \ell = 0$, this corresponds roughly to a torque of $10^{-3}$ times the maximum torque at perigee. Unless otherwise indicated, the simulation is terminated at $D=10 r_p$ as well.

With the simulation inputs specified, the equations of motion were integrated via the Runge-Kutta fourth order method, with a variable time step
\begin{equation}
  dt = dt_\text{min} + 10^{-3}(dt_\text{max} - dt_\text{min}) \brackets{\parens{\frac{D}{r_p}}^3 - 1}.
\end{equation}
The parameters $dt_\text{max}$ and $dt_\text{min}$ (20 and 10 seconds respectively) were chosen such that the error was $\sim$100 times the floating point error, and that neighbouring values of $K_{\ell m}$ yielded significantly different spin pole data compared to floating point error. The data used to choose this $dt$ was obtained using the reference asteroid configuration, described in appendix \ref{app:reference-configs}.

\section{Injection-retrieval tests}
\label{sec:injection-retrieval}

In this section, we discuss how we used the simulation described in the previous section to generate synthetic data and extract density moment posterior distributions from the data.

\subsection{Fit process}
\label{sec:fit}
Given synthetic data, an Affine Invariant Markov Chain Monte Carlo (MCMC) Ensemble sampler was used to generate posterior distributions from flat priors. We use the Python implementation \texttt{emcee} \cite{foreman2013emcee}. Our parameters were $\gamma_0$, $K_{20}$, $K_{22}$, and $K_{3m}$ (10 in total), and were bounded by $|\gamma_0| < \pi/4$, and bounds on $K_{2 m}$ given in equation \ref{eqn:parameter-bounds}. Note that $\gamma_0$ is degenerate with $\gamma_0 + \pi/2$ since this is equivalent to re-labeling $\unit y$ as $\unit x$ and $\unit x$ as $-\unit y$. The other bounds were $|K_{3m}| < 1$.

The MCMC was determined to converge when the fractional change in autocorrelation time (computed every 100 iterations) was one percent, and the number of iterations computed so far was more than 100 times the autocorrelation time. The MCMC fit also was set to terminate if more than $10^5$ iterations were run, but this only occurred for fits in which the data was too low quality to resolve the parameters. $~10^4$ iterations was often required, which generally consumed about 7 hours of computation time on a super computer running 16 threads on 8 cores.

Before the MCMC was run, local minima in the likelihood were found via the Nelder-Mead algorithm implemented in \texttt{scipy} \cite{Gao2012}. It was found that only one local minimum existed, except when $K_{22}=0$ in which case rotational symmetry caused multiple values of $\gamma_0$ to be degenerate. Walkers were initialized near this local minimum, distributed with probability distributed as a Gaussian approximation of the likelihood, as determined via the inverse Hessian of the likelihood at the minimum. Due to the high sensitivity of the angular velocity data to density moments, the minimization procedure sometimes failed to isolate the minimum likelihood. Therefore, a simpler simulation without the $K_{3m}$ terms of equation \ref{eqn:tidal-torque} was first used to minimize likelihood as a function of $\gamma_0$ and $K_{2m}$, and then the full simulation was used to find $K_{3m}$, with $\gamma_0$ and $K_{2m}$ fixed. This tiered minimization process motivates us to call $\gamma_0$ and $K_{2m}$ the ``first-order parameters'' and $K_{3m}$ the ``second-order parameters.''

We further subdivided the minimization process by first minimizing with respect to data truncated at the point after perigee where $D=fr_p$ for some manually-set fraction $f$. In practice, $f=2$ often led to success. The minimum was then further refined by minimizing based on the full data, with the previous minimum as the initial estimate.

\subsection{Uncertainty model}
\label{sec:uncertainty}

We model uncertainty in each asteroid spin vector $\bm \Omega$ in the as uncorrelated with other spin vectors, and we model uncertainty in the orientation and in the period as also uncorrelated. Consider a true spin vector $\bm \Omega^*$. For the sake of description, we work in coordinates in which $\bm \Omega^* \parallel \unit z$. Then, expressing the observed spin vector $\bm \Omega$ in spherical coordinates, we draw the polar angle from a normal distribution with standard deviation $\sigma_\theta$ centred on zero and the azimuthal angle from a uniform distribution. We also draw the ratio $\rho=\Omega/\Omega^*$ from a log-normal distribution centred on one, with width $\sigma_\rho$. We generally represent $\sigma_\theta > \sigma_\rho$ under the assumption that period is better constrained by light-curve data than spin pole. Explicitly, the probability density function (PDF) of $\rho$ is 
\begin{equation}
  P(\rho) = \frac{1}{\rho\sqrt{2\pi \sigma_\rho^2}} \exp\parens{-\frac{\ln^2\rho}{2\sigma_\rho^2}}.
\end{equation}
See figure \ref{fig:uncertainty-model} for an illustration of the uncertainty model shown in the inertial frame.

\begin{figure}
  \centering
  \begin{tikzpicture}
  \draw[-{Latex[length=3mm]}] (0, 0) -- (-4, 0) node[anchor=east] {$\unit X$};
  \draw[-{Latex[length=3mm]}] (0, 0) -- (1, -1.5) node[anchor=west] {$\unit Y$};
  \draw[-{Latex[length=3mm]}] (0, 0) -- (0, 4) node[anchor=south] {$\unit Z$};

  \draw[line width=0.5mm, -{Latex[length=3mm]}] (0, 0) -- (-3, 3) node[anchor=south] {$\bm \Omega^*$};
  \draw[dashed] (-3, -2) -- (-3, 3);
  \draw[dashed] (-3, -2) -- (0, 0);

  \draw[-{Latex[length=3mm]}] (0, 0) -- (-1.25, 3.35) node[anchor=south west] {$\bm \Omega$};
  \draw[rotate around={-50:(-3,3)}] (-3,3) ellipse (1.4 and 2);
  \draw (-3, 3) arc (110:93:6);
  \draw (-2, 3.5) node[anchor=center] {$\theta$};
  \draw (-1, 1.8) node[anchor=center] {$\rho \Omega^*$};

  \end{tikzpicture}
  \caption{Diagram in the inertial frame of the uncertainty model used to define the probability that the true spin vector $\bm \Omega^*$ should be observed as $\bm \Omega$. The parameter $\theta$ is drawn from a Gaussian with width $\sigma_\theta$, and $\rho$ is drawn from a log normal distribution with width $\sigma_\rho$.}
  \label{fig:uncertainty-model}
\end{figure}

The log likelihood resulting from this uncertainty model is (excluding additive constants)
\begin{equation}
  \begin{split}
  \ln \mathcal{L} = -\frac{1}{2}\sum_{i = 0}\Bigg[&\frac{\cos^{-1} (\bm \Omega_i^* \cdot \bm \Omega_i/(\Omega_i^* \Omega_i))^2}{\sigma_\theta^2}\\
  +&\frac{\ln \parens{\Omega_i /\Omega_i^*}^2}{\sigma_\rho^2} + 2\ln\frac{\Omega_i}{\Omega_i^*}\Bigg].
  \end{split}
  \label{eqn:log-likelihood}
\end{equation}
where $\Omega_i$ is the $i$th spin vector in the data set.

In figure \ref{fig:example-data}, we present example spin data generated via this simulation. A population of one thousand asteroids with identical initial conditions except for $\gamma_0$, $K_{20}$, and $K_{22}$ were simulated on a close Earth encounter. The exact parameters used were the symmetric and asymmetric cases described in appendix \ref{app:reference-configs}. Bands containing 68.3\%, 95.5\%, and 99.7\% of the population's spin are shown, as is the spin of the reference asteroids in black.

\begin{figure*}
  \centering
  \includegraphics[width=0.7\linewidth]{figs/nominal-data-sym.pdf}
  \includegraphics[width=0.7\linewidth]{figs/nominal-data-asym.pdf}
  \caption{Angular velocity data simulated for the symmetric (above) and asymmetric (below) reference asteroids. The true angular velocity evolution is shown as a black line. Also plotted is the deviation of the data for posterior-PDF-distributed perturbations to the asteroid shape (bands). Bands contain 68.3\%, 95.5\%, and 99.7\% of the 1000 simulations run.}
  \label{fig:example-data}
\end{figure*}

To choose the initial values of $\gamma_0$, $K_{20}$, and $K_{22}$, spin data for the reference asteroids of appendix \ref{app:reference-configs} was first generated. Then $\gamma_0$, $K_{20}$, and $K_{22}$ were re-extracted via the fit described in section \ref{sec:fit}. The population's values for these parameters were posterior-distributed. However, the posterior distribution was widened by a factor of 1000 to make the band widths visible. Therefore, the scale of the bands in figure \ref{fig:example-data} have little meaning in an absolute sense, but they are meaningful when comparing two bands or two times in the same band.

The figure illustrates the sensitivity of spin data to asteroid density moments and $\gamma_0$; Before perigee, all asteroids had similar angular velocities, but after perigee the angular velocities of the population diverged. The oscillation periods observable during the torque-free precession phase of the spin data (about 5 hours after perigee) vary within the population, showing that they can be used to constrain $K_{22}$ and $K_{20}$.

\section{Results}
\label{sec:results}

In this section, we assess the sensitivity of the posterior distributions for the first-order parameters ($\gamma_0$, $K_{20}$, and $K_{22}$), and the second-order parameters ($K_{3m}$) to many possible close encounter configurations. Specifically, we test sensitivity to the encounter's orbital parameters, the degree of observational uncertainty, the asteroid's radius, the cadence of observation, the amount of data collected at perigee, the direction of the initial spin pole, the asteroid period, and the central body oblateness. These effects are presented roughly in order of how much they affect the parameter precision, from highest to lowest.

Parameter precision of 1$\sigma$ is defined such that 68.27\% of the posterior distribution lies within 1$\sigma$ of the mean of the posterior distribution. 2$\sigma$ is defined likewise for 95.45\% of the posterior. The posteriors are usually roughly Gaussian, so that the upper and lower error bars are usually roughly equal and 1$\sigma$ is approximately the standard deviation of the posterior.

In all cases, we use the configuration of the asymmetric reference asteroid (appendix \ref{app:reference-configs}) unless otherwise stated. We further present a test of how precision depends on cadence, asteroid period, and the duration of the encounter in appendix \ref{app:cadence-tests}, and we compare precision for a Jupiter and an Earth encounter in appendix \ref{app:jupiter-earth}.

\subsection{Orbital elements}
\label{sec:scan-orbit}
A Keplerian orbit is completely described by five parameters, but three describe the orbit's orientation with respect to the central body. The orbit can be rotated to the $\unit X \unit Y$-plane, as mandated by our coordinate definitions, by changing the density moments of the central body. Since $J_{00}$ is unchanged by this rotation and $J_{1m}=0$, the orbit orientation is irrelevant until the $J_{2m}$ terms of equation \ref{eqn:tidal-torque}, and we do not investigate them here.

We parametrize the shape of the orbit by the perigee distance $r_p$ and excess velocity $v_\infty$. Fits of the type described in section \ref{sec:fit} were run for many values of $r_p$ and $v_\infty$ and the 1 and 2$\sigma$ confidence intervals are displayed in figures \ref{fig:scan-vex} for $v_\infty$ and \ref{fig:scan-perigee} for $r_p$.

\begin{figure}
  \centering
  \includegraphics[height=0.89\textheight]{figs/scan-vex.pdf}
  \caption{1 and 2$\sigma$ confidence intervals for the first-order parameter posteriors (\textit{top}) and second-order parameters (\textit{bottom}) as a function of orbital excess velocity $v_\infty$. The vertical dashed line indicates the reference asteroid value of 6 km s$^{-1}$. The red vertical lines indicate when $\sigma = 0.01$.}
  \label{fig:scan-vex}
\end{figure}

Figure \ref{fig:scan-vex} demonstrates that parameter precision does not depend strongly on excess velocity, aside from a slight trend especially in the higher order parameters for uncertainty to increase with $v_\infty$. This is likely due to the fact that larger $v_\infty$ leads to a faster and flatter orbit with less time spent close to the planet, where tidal torque is strongest. There are also smaller-scale oscillations in the uncertainty, due to the orientation of the asteroid at perigee varying. The asteroid is always simulated to start at the same orientation, but increasing $v_\infty$ decreases the time to perigee, so that the asteroid enters this region of high torque at different orientations depending on $v_\infty$. This effect explains why these small-scale oscillations have the same period for all parameters. Note that these oscillations are sometimes large enough to raise $\sigma > 0.01$, shown by the red vertical line.

\begin{figure}
  \centering
  \includegraphics[height=0.89\textheight]{figs/scan-perigee.pdf}
  \caption{1 and 2$\sigma$ confidence intervals for the first-order parameter posteriors (\textit{top}) and second-order parameters (\textit{bottom}) as a function of perigee distance $r_p$. The vertical dashed line indicates the reference asteroid value of 5 Earth radii. The dotted curve indicates a power-law fit (see text). The red vertical lines indicate when $\sigma = 0.01$.}
  \label{fig:scan-perigee}
\end{figure}

Figure \ref{fig:scan-perigee} shows much stronger dependence of parameter uncertainty on perigee distance, as expected by the factor of $(a_\mathcal{A}/D)^{\ell'}$ present in equation \ref{eqn:tidal-torque} and mentioned in section \ref{sec:tidal-torque}. For $r_p \approx 6$ or 7 Earth radii, $\sigma(K_{30})$ reaches the reference 0.01 limit. At slightly higher $r_p$, $K_{31}$ and $K_{32}$ become similarly unresolved, and at $r_p \approx 20$ Earth radii, $K_{33}$, the last second-order component, becomes unresolved. Even before this, at $r_p \approx 10$ Earth radii, the most uncertain parameter $K_{30}$ fills the prior distribution with uncertainty ranging from -1 to 1, visible by the sudden cut-off in uncertainty increase and the discontinuity of the $\sigma$ curve there. The location of these limits will change if different encounter properties are used, but this is still an illustrative example.

\begin{table}
  \centering
  \begin{tabular}{l|c}
    \hline
    Parameter & $\alpha$ \\ \hline
    $\gamma_0$ & $2.05$ \\
    $K_{22}$ & $5.47$ \\
    $K_{20}$ & $5.47$ \\ \hline
    $\Re K_{33}$ & $3.35$ \\
    $\Im K_{33}$ & $3.37$ \\
    $\Re K_{32}$ & $3.05$ \\
    $\Im K_{32}$ & $3.27$ \\
    $\Re K_{31}$ & $3.53$ \\
    $\Im K_{31}$ & $4.02$ \\
    $K_{30}$ & $5.75$ \\ \hline
  \end{tabular}
  \caption{Power law slope values for the dependence of parameter uncertainty on perigee distance $r_p$. Slope is defined by $\sigma \propto r_p^\alpha$.}
  \label{tab:scan-perigee-alpha}
\end{table}

Fitted to each of the curves in figure \ref{fig:scan-perigee} are power law uncertainties $\sigma \propto r_p^\alpha$. These fits were performed via the method of least squares, and all data with $\sigma > 0.7$ was removed due to its sensitivity to the arbitrarily-chosen boundary of the prior. The values of $\alpha$ are shown in table \ref{tab:scan-perigee-alpha}. These slope values express how much each parameter is dependent on $r_p$. It is observed that $\gamma_0$ is least dependent on $r_p$, with $\sigma(\gamma_0) \sim r_p^2$. The other two first-order parameters are much more strongly dependent, with $\sigma \sim r_p^{5.5}$. The second-order parameters have milder slopes between 3 and 4 (except for $K_{30}$) which is fortunate from the perspective of making precise observations, since lower $\alpha$ renders larger values of $r_p$ accessible to measuring $K_{3m}$.

The axes of figure \ref{fig:scan-perigee} show that parameters with large $m$ are more precisely determined than parameters with small $m$, as can be seen by comparing $K_{22}$ to $K_{20}$ and comparing $K_{33}$ to other $K_{3m}$ values. Large $m$ moments correspond to moments that control higher frequency fluctuations in density at the asteroid equator. This pattern of large $m$ corresponding to low uncertainty is a general trend and will be seen in the following sections as well.

The very strong dependence of $\sigma$ on $r_p$ makes this analysis only to extract second-order moments on close encounters. Fortunately, in the case of Earth, these encounters are also likely to have the best associated observational uncertainty when above the horizon due to their proximity. The first-order moments can still be extracted at much larger perigee distances in our model.

\subsection{Observational Uncertainty}
\label{sec:scan-uncertainty}
There are two parameters, $\sigma_\theta$ and $\sigma_\rho$, which govern the observational uncertainty of the data set (defined in section \ref{sec:uncertainty}). Rather than explore the full space spanned by these two values, we measure how parameter uncertainties depend on the product of uncertainties $\sigma_\theta\sigma_\rho$ (with the radio fixed), and the ratio $\sigma_\rho / \sigma_\theta$ (with the product fixed).

We choose these metrics because we generally expect that the parameter uncertainty $\sigma$ be proportional to the observational uncertainty, but whether the dependence is stronger on $\sigma_\theta$ or $\sigma_\rho$ is not immediately clear. We get around this problem by varying $\sigma_\theta$ and $\sigma_\rho$ together and fixing their ratio, and measuring the posterior uncertainty $\sigma$, with $\sigma/\sigma_\theta$ shown in figure \ref{fig:scan-product}. Indeed, we find that $\sigma \propto \sigma_\theta$ almost exactly, and since $\sigma_\rho / \sigma_\theta$ is fixed, we also have $\sigma \propto \sigma_\rho$. For large $\sigma_\theta \sigma_\rho$, the proportionality fails, but this is because $\sigma(K_{30}) \approx 1$ which fills the prior. Uncertainty cannot increase beyond this value.

\begin{figure}
  \centering
  \includegraphics[height=0.89\textheight]{figs/scan-product.pdf}
  \caption{1 and 2$\sigma$ confidence intervals divided by $\sigma_\theta$ for the first-order parameter posteriors (\textit{top}) and second-order parameters (\textit{bottom}) as a function of observational uncertainty product $\sigma_\theta \sigma_\rho$. The vertical dashed line indicates the reference asteroid value of $10^{-9}$. The red vertical lines indicate when $\sigma =0.01$.}
  \label{fig:scan-product}
\end{figure}

We also investigate the dependence of posterior uncertainty on $\sigma_\rho / \sigma_\theta$ with $\sigma_\theta \sigma_\rho$ fixed in figure \ref{fig:scan-ratio}. If we simply had $\sigma \propto \sigma_\theta \sigma_\rho$, then $\sigma$ would have no dependence on $\sigma_\rho / \sigma_\theta$. Any deviation from constant $\sigma$ shown in the figure therefore reveals some additional dependence in the model on one of the observational uncertainties.

\begin{figure}
  \centering
  \includegraphics[height=0.89\textheight]{figs/scan-ratio.pdf}
  \caption{1 and 2$\sigma$ confidence intervals for the first-order parameter posteriors (\textit{top}) and second-order parameters (\textit{bottom}) as a function of observational uncertainty ratios $\sigma_\rho / \sigma_\theta$. The reference asteroid value is $\sigma_\rho/ \sigma_\theta =10^{-5}$. The red vertical lines indicate when $\sigma = 0.01$. (Lines are not shown for $K_{3m}$, $m < 3$ because they coincide with the vertical axis.)}
  \label{fig:scan-ratio}
\end{figure}

Indeed, figure  \ref{fig:scan-ratio} shows increased uncertainty when $\sigma_\rho/\sigma_\theta$ is large, so that $\sigma$ depends more on $\sigma_\rho$ than on $\sigma_\theta$. We summarize this pattern by stating that, if the observer had to choose between better precision on the period or on the spin pole of the data, they should choose period. This is fortunate for observers since, \jtd{I want to say that precision on the period is better constrained by light curve analysis, but I do not know enough about light curve analysis to be sure.}

Another conclusion that can be drawn from figure \ref{fig:scan-ratio} is that $\sigma(\sigma_\rho/\sigma_\theta)$ is proportional for all $K_{\ell m}$. The proportionality constant shows the same dependence on $\ell$ and $m$ mentioned previously. The proportionality is broken near $\sigma \approx \pm 1$, where $\sigma$ fills the prior.

>>>>

@@

\section{Tidal torque \& Equations of motion}
\label{app:eom}

In this appendix, we derive the equations of motion used to simulate the asteroid angular velocity during the encounter. In particular, we describe our coordinates (section \ref{sec:coordinates}) for an encountering asteroid's position and orientation, and we parametrize its density distribution via its ``density moments'' (section \ref{sec:moments}). Then we derive an arbitrary-order equation for tidal torque (section \ref{sec:tidal-torque}) and write the equations of motion for the system (section \ref{sec:eom}).

@@

We define the density moments of an asteroid in equation \ref{eqn:klm}.
Note that these are complex. Here, $\mathcal{A}$ indicates the volume of the asteroid, $\mu_\mathcal{A}$ is the mass of the asteroid, $\rho_\mathcal{A}(\bm r)$ is the density distribution, and $a_\mathcal{A}$ is a real length scale defined in equation \ref{eqn:am}. This length can be thought of as akin to the radius of the asteroid (although a spherical, uniform-density asteroid has radius $a_\mathcal{A}\sqrt{5/3} $). Both equations \ref{eqn:klm} and \ref{eqn:am} should be computed in the body-fixed axes.

@@

Throughout this paper, we assume that the asteroid remains rigid throughout the encounter. We also assume no third-body perturbations from other Solar System objects. (Actually, third-body perturbing objects are allowed if they are closer to the central body's centre of mass than the asteroid perigee distance. Then, their density moments can be included in the density moments of the central body and this derivation can still be used.) For the sake of simplicity, we also assume that the density moments of the central body are known and do not evolve with time (i.e., the central body's rotation is marginal compared to the timescale of the encounter).

@@

The tidal torque experienced by the asteroid (in the body-fixed frame) is given by
\begin{equation}
  \bm{\tau}(\bm u) = \int_\mathcal{A} d^3 u \rho_\mathcal{A}(\bm u) (\bm u \times (-\nabla_{\bm u} V(\bm u)))
\end{equation}
where $\rho_\mathcal{A}$ is the density distribution of the asteroid and $\mathcal{A}$ indicates the volume of the asteroid. Making use of one more identity concerning the derivatives of spherical harmonics:
\begin{equation}
  \begin{split}
  \bm u \times \nabla R_{\ell m}(\bm u)=&\frac{1}{2}\Big[(i\unit x - \unit y)(\ell-m+1)R_{\ell,m-1}(\bm u)\\
  &+(i\unit x+\unit y)(\ell+m+1)R_{\ell,m+1}(\bm u)\\
  & +2im\unit z R_{\ell m}(\bm u)\Big],
  \end{split}
\end{equation}
tidal torque can now be expressed as a function only of $J_{\ell m}$, $K_{\ell m}$, $a_\mathcal{A}$, $a_\mathcal{B}$, and the asteroid orientation and position. This equation is given explicitly as equation \ref{eqn:tidal-torque}. Some $K_{\ell m}$ terms are included in this equation with $|m|>\ell$; these should all be taken to be zero.

@@
